---
title: "Testing Metrics"
output:
  html_document:
    css: "site_themes.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(reticulate)
library(vegawidget)
```

```{python include=FALSE}
import csv
import numpy as np

import pandas as pd
import altair as alt
```



```{python include=FALSE}

def read_csv(filepath):
    # This function will load a csv when passed
    # a filepath and returns a numpy array 
    with open(filepath,"r") as f:
        reader=csv.reader(f)
        return np.array([x for x in reader])

def get_column_index(labels,col_str):
    # returns the index of a column
    # pass the column as a string with the 
    # current label state
    return np.where(labels==col_str)[0][0]

rdt_mets=read_csv('/Users/jason/OneDrive - UW/Documents/research_projects/202109_dedx/data/rdt_metrics2.csv')
rdt_mets[0][0]="PRODUCT"

rdtm_labs=rdt_mets[0]

rdtm_data=rdt_mets[1:,:]

rdtm_df=pd.DataFrame(rdtm_data,columns=rdtm_labs)

rdtm_df[['CLINICAL_SENSITIVITY_mean','CLINICAL_SENSITIVITY_CI_low','CLINICAL_SENSITIVITY_CI_high','CLINICAL_SPECIFICITY_mean','CLINICAL_SPECIFICITY_CI_low','CLINICAL_SPECIFICITY_CI_high']]=rdtm_df[['CLINICAL_SENSITIVITY_mean','CLINICAL_SENSITIVITY_CI_low','CLINICAL_SENSITIVITY_CI_high','CLINICAL_SPECIFICITY_mean','CLINICAL_SPECIFICITY_CI_low','CLINICAL_SPECIFICITY_CI_high']].apply(pd.to_numeric, errors='coerce', axis=1).apply(lambda x: x/100)

# rdtm_dfs=rdtm_df.sort_values("CLINICAL_SENSITIVITY_mean",ascending=False)


def filter_metrics(sort_term):
    return rdtm_df[rdtm_df['READ']==sort_term].sort_values("CLINICAL_SENSITIVITY_mean",ascending=False)

def sens_plot(sort_term,subtitle_text):
    filter_df=filter_metrics(sort_term).sort_values("CLINICAL_SENSITIVITY_mean",ascending=False)
    sens_base = alt.Chart(filter_df).encode(
        x=alt.X('PRODUCT', sort=filter_df['PRODUCT'].to_list(),axis=alt.Axis(title="Antigen Diagnostic Tests for SARS-CoV-2")),
        color=alt.Color('PRODUCT',legend=None)
    ).properties(
        title={
            'text':'Reported Clinical Sensitivity', 
            'fontSize':24,
            'subtitle':subtitle_text,
            'subtitleFontSize':18},
        width=400
    )
    sens_rule = sens_base.mark_rule().encode(
        alt.Y(
            'CLINICAL_SENSITIVITY_CI_low:Q',
            title='Clinical Sensitivity',
            scale=alt.Scale(zero=False),
        ),
        alt.Y2('CLINICAL_SENSITIVITY_CI_high:Q')
    )

    sens_bar = sens_base.mark_bar(opacity=0.5).encode(
        alt.Y('CLINICAL_SENSITIVITY_CI_low:Q'),
        alt.Y2('CLINICAL_SENSITIVITY_CI_high:Q'),
        tooltip=[alt.Tooltip('PRODUCT:N',title='Antigen Test'),alt.Tooltip('CLINICAL_SENSITIVITY_mean:Q',format=".3p",title='Mean'),alt.Tooltip('CLINICAL_SENSITIVITY_CI_low:Q',format=".3p", title="Low"),alt.Tooltip('CLINICAL_SENSITIVITY_CI_high:Q',format=".3p",title='Hi')]
    )
    sens_points = sens_base.mark_point().encode(
        y=alt.Y('CLINICAL_SENSITIVITY_mean:Q'),
        tooltip=[alt.Tooltip('PRODUCT:N',title='Antigen Test'),alt.Tooltip('CLINICAL_SENSITIVITY_mean:Q',format=".3p",title='Mean'),alt.Tooltip('CLINICAL_SENSITIVITY_CI_low:Q',format=".3p", title="Low"),alt.Tooltip('CLINICAL_SENSITIVITY_CI_high:Q',format=".3p",title='Hi')] 
        #color=alt.Color('PRODUCT',legend=None)
    )


    return alt.layer(sens_bar,sens_rule,sens_points)


def spec_plot(sort_term,subtitle_text):
    filter_df=filter_metrics(sort_term).sort_values("CLINICAL_SPECIFICITY_mean",ascending=False)
    sens_base = alt.Chart(filter_df).encode(
        x=alt.X('PRODUCT', sort=filter_df['PRODUCT'].to_list(),axis=alt.Axis(title="Antigen Diagnostic Tests for SARS-CoV-2")),
        color=alt.Color('PRODUCT',legend=None)
    ).properties(
        title={
            'text':'Reported Clinical Specificity', 
            'fontSize':24,
            'subtitle':subtitle_text,
            'subtitleFontSize':18},
        width=400
    )
    sens_rule = sens_base.mark_rule().encode(
        alt.Y(
            'CLINICAL_SPECIFICITY_CI_low:Q',
            title='Clinical Sensitivity',
            scale=alt.Scale(zero=False),
        ),
        alt.Y2('CLINICAL_SPECIFICITY_CI_high:Q')
    )

    sens_bar = sens_base.mark_bar(opacity=0.5).encode(
        alt.Y('CLINICAL_SPECIFICITY_CI_low:Q'),
        alt.Y2('CLINICAL_SPECIFICITY_CI_high:Q'),
        tooltip=[alt.Tooltip('PRODUCT:N',title='Antigen Test'),alt.Tooltip('CLINICAL_SPECIFICITY_mean:Q',format=".3p",title='Mean'),alt.Tooltip('CLINICAL_SPECIFICITY_CI_low:Q',format=".3p", title="Low"),alt.Tooltip('CLINICAL_SPECIFICITY_CI_high:Q',format=".3p",title='Hi')]
    )
    sens_points = sens_base.mark_point().encode(
        y=alt.Y('CLINICAL_SPECIFICITY_mean:Q'),
        tooltip=[alt.Tooltip('PRODUCT:N',title='Antigen Test'),alt.Tooltip('CLINICAL_SPECIFICITY_mean:Q',format=".3p",title='Mean'),alt.Tooltip('CLINICAL_SPECIFICITY_CI_low:Q',format=".3p", title="Low"),alt.Tooltip('CLINICAL_SPECIFICITY_CI_high:Q',format=".3p",title='High')] 
        #color=alt.Color('PRODUCT',legend=None)
    )


    return alt.layer(sens_bar,sens_rule,sens_points)


test_options=["visual read","instrument read","interpreted results presented"]

viz_rd_sens_viz=sens_plot(test_options[0], "Visual Read Diagnostic Tests")


instr_rd_sens_viz=sens_plot(test_options[1],"Instrument Read Diagnostic Tests")


poc_pcr_sens_viz=sens_plot(test_options[2],"POC PCR Diagnostic Tests")


viz_rd_spec_viz=spec_plot(test_options[0], "Visual Read Diagnostic Tests")


instr_rd_spec_viz=spec_plot(test_options[1],"Instrument Read Diagnostic Tests")


poc_pcr_spec_viz=spec_plot(test_options[2],"POC PCR Diagnostic Tests")


viz_rd_df,instr_rd_df,poc_pcr_df=filter_metrics(test_options[0]),filter_metrics(test_options[1]),filter_metrics(test_options[2])


agg_spec=alt.Chart(rdtm_df).mark_boxplot(extent='min-max',size=69).encode(
        x=alt.X('READ:N',sort=test_options,axis=alt.Axis(labelAngle=20)),
        y=alt.Y('CLINICAL_SPECIFICITY_mean',title='Specificity',scale=alt.Scale(domain=(0.95,1))),
        color="READ:N",
    ).properties(
            title={
                'text':str('Reported Clinical Specificity'), 
                'fontSize':24,
                'subtitle':'Aggregated Stats from Each Testing Group',
                'subtitleFontSize':18},
            width=600
        )

agg_sens=alt.Chart(rdtm_df).mark_boxplot(extent='min-max',size=69).encode(
        x=alt.X('READ:N',sort=test_options,axis=alt.Axis(labelAngle=20)),
        y=alt.Y('CLINICAL_SENSITIVITY_mean',title='Sensitivity',scale=alt.Scale(domain=(0.8,1))),
        color="READ:N",
    ).properties(
            title={
                'text':str('Reported Clinical Sensitivity'), 
                'fontSize':24,
                'subtitle':'Aggregated Stats from Each Testing Group',
                'subtitleFontSize':18},
            width=600
        )

```



## Diagnostic Tests for SARS-CoV-2 {.tabset}

### Individual Tests

```{python echo=FALSE}
ind_chart=(viz_rd_sens_viz|viz_rd_spec_viz)&(instr_rd_sens_viz|instr_rd_spec_viz)&(poc_pcr_sens_viz|poc_pcr_spec_viz)

ichart=ind_chart.to_json()
```

```{r echo=FALSE}
as_vegaspec(py$ichart)
```

- These charts show the individual tests in each of our assigned classes. The bar indicates the 95% confidence intervals.

- The data was compiled in <a href="https://pubs.acs.org/doi/pdf/10.1021/acsinfecdis.1c00268">this</a> article. It was collect from the kit handbooks from the EUA-FDA SARS-CoV-2 diagnostic tests (Pub Date: September 27, 2021)


### Compiled Metrics

```{python echo=FALSE}
agg_chart=agg_sens&agg_spec
achart=agg_chart.to_json()
```

```{r echo=FALSE}
as_vegaspec(py$achart)
```

- Here are the distributions of each testing class. "interpreted results presented" are POC NAATs

### Table

```{r include=FALSE}
library(kableExtra)
library(tidyverse)
```


```{r echo=FALSE}
tab <- py$rdtm_df

tab2 <- tab%>%select(-c(starts_with("LOD"),"NOTES"))

tab2 %>%
  kbl(col.names = c('Product','Test Type','Test Class (Internal)','Mean','Low','High','Mean','Low','High')) %>%
  kable_styling(c("striped", "hover"))%>%
  column_spec(4, color = spec_color(tab2$CLINICAL_SENSITIVITY_mean,option="C", end = 0.7)) %>%
  add_header_above(c(" " = 3, "Reported Sensitivity" = 3, "Reported Specificity" = 3))%>%
  scroll_box(width = "100%", height = "90%")

```

